---
title: "R Basic Notebook"
output: html_notebook
---

# Introduction to R programming 

Assignment operation - Create a list in R 
```{r}
rand_var <- c(1,2,3,4)
rand_var

rand_var[1]
```

Reading a csv file 
```{r}
sample_data <- read.csv("./data/simple.csv", head = TRUE, sep = ",")
sample_data
```


Descriptove stats about the dataset 
```{r}
summary(sample_data)
```

To get more information about the function of interest 
```{r}
help(read.csv)
```

dir - To list all the files in a directory 

getwd - Gives you the path location of where exactly the files are stored 

```{r}
dir()

getwd()
```

Accessing the data frame columns 

```{r}
sample_data$trial

sample_data$mass

sample_data$velocity
```

Get the column names 

```{r}
names(sample_data)
```

attributes - Gives information about the names of the column, class and number of values under each column 

```{r}
attributes(sample_data)
```

# Variable Types 

## Numbers 

```{r}
a <- 3 
print(a)
b <- sqrt(a*a+3)
print(b)
ls()
```

```{r}
a <- c(1,2,3,4,5)
print(mean(a))
print(var(a))
```

```{r}
a[0]
```

```{r}
a[6]
```

```{r}
a <- numeric(10)
a
```

Determine the data type used for a variable

```{r}
typeof(a)
```

## Strings 

```{r}
b <- c("hello","there")
b
```

```{r}
a <- "hello"
a
```

## Factors

```{r}
summary(sample_data$trial)
```

```{r}
sample_data$velocity <- factor(sample_data$velocity)
sample_data$velocity
```

```{r}
summary(sample_data$velocity)
```

```{r}
levels(sample_data$velocity)
```

## Dataframes 

```{r}
a <- c(1,2,3,4)
b <- c(2,4,6,8)
levels <- factor(c("A","B","A","B"))
bubba <- data.frame(first=a,second=b,f=levels)

bubba
summary(bubba)
```

## Logical opearations

```{r}
a = TRUE
typeof(a)
```

```{r}
a = c(TRUE,FALSE)
b = c(FALSE,FALSE)
a|b
```

## Tables 

One way tables

```{r}
a <- factor(c("A","A","B","A","B","B","C","A","C"))
results <- table(a)
results
summary(results)
attributes(results)
```

```{r}
occur <- matrix(c(4,3,2),ncol=3,byrow=TRUE)
occur
```

Two way tables 

```{r}
a <- c("Sometimes","Sometimes","Never","Always","Always","Sometimes","Sometimes","Never")
b <- c("Maybe","Maybe","Yes","Maybe","Maybe","No","Yes","No")
results <- table(a,b)
results
```

# Numerical operations

## Basic operations

```{r}
a <- c(1,2,3,4)
a+5
a-5
a*2
a/2
```

```{r}
sqrt(a)
exp(a)
log(a)
exp(log(a))
```

```{r}
a <- c(1,-2,3,-4)
b <- c(-1,2,-3,4)
min(a,b)
pmin(a,b)
```

Numerical Description

```{r}
mean(sample_data$mass)
median(sample_data$mass)
quantile(sample_data$mass)
min(sample_data$mass)
max(sample_data$mass)
var(sample_data$mass)
sd(sample_data$mass)
```

Operations on vector

```{r}
a = c(2,4,6,3,1,5)
b = sort(a)
b
c = sort(a,decreasing = TRUE)
c
```

# Probability Distribution

Normal Distribution

The first function we look at it is dnorm. Given a set of values it returns the height of the probability distribution at each point. If you only give the points it assumes you want to use a mean of zero and standard deviation of one. 

```{r}
dnorm(0)
dnorm(0)*sqrt(2*pi)
dnorm(0,mean=4)
dnorm(0,mean=4,sd=10)
v <- c(0,1,2)
dnorm(v)
x <- seq(-20,20,by=.1)
y <- dnorm(x)
plot(x,y)
y <- dnorm(x,mean=2.5,sd=0.1)
plot(x,y)
```

The second function we examine is pnorm. Given a number or a list it computes the probability that a normally distributed random number will be less than that number. This function also goes by the rather ominous title of the ???Cumulative Distribution Function.???

```{r}
pnorm(0)
pnorm(1)
pnorm(0,mean=2)
pnorm(0,mean=2,sd=3)

v <- c(0,1,2)
pnorm(v)

x <- seq(-20,20,by=.1)

y <- pnorm(x)
plot(x,y)
y <- pnorm(x,mean=3,sd=4)
plot(x,y)
```

The next function we look at is qnorm which is the inverse of pnorm. The idea behind qnorm is that you give it a probability, and it returns the number whose cumulative distribution matches the probability

```{r}
x <- seq(0,1,by=.05)
y <- qnorm(x)
plot(x,y)
y <- qnorm(x,mean=3,sd=2)
plot(x,y)
y <- qnorm(x,mean=3,sd=0.1)
plot(x,y)
```

The last function we examine is the rnorm function which can generate random numbers whose distribution is normal.

```{r}
rnorm(4)
rnorm(4,mean=3)
rnorm(4,mean=3,sd=3)

y <- rnorm(200)
hist(y)

y <- rnorm(200,mean=-2)
hist(y)

y <- rnorm(200,mean=-2,sd=4)
hist(y)

qqnorm(y)
qqline(y)
```

## T Distribution

```{r}
x <- seq(-20,20,by=.5)
y <- dt(x,df=10)

plot(x,y)

y <- dt(x,df=50)
plot(x,y)

```

## Binomial Distribution 

```{r}
x <- seq(0,50,by=1)
y <- dbinom(x,50,0.2)

plot(x,y)
y <- dbinom(x,50,0.6)
plot(x,y)

x <- seq(0,100,by=1)
y <- dbinom(x,100,0.6)
plot(x,y)
```

## Chi-Squared Distriution 

```{r}
x <- seq(-20,20,by=.5)
y <- dchisq(x,df=10)
plot(x,y)
y <- dchisq(x,df=12)
plot(x,y)
```

# Basic Plots

```{r}
hist(sample_data$mass,main="Distribution of Mass",xlab="Mass")
boxplot(sample_data$mass, main="Outliers - Mass", ylab="Mass")
plot(sample_data$mass,sample_data$velocity)
```

```{r}
x <- rnorm(10,sd=5,mean=20)
y <- 2.5*x - 1.0 + rnorm(10,sd=9,mean=0)
cor(x,y)

x1 <- runif(8,15,25)
y1 <- 2.5*x1 - 1.0 + runif(8,-6,6)

x2 <- runif(8,15,25)
y2 <- 2.5*x2 - 1.0 + runif(8,-6,6)

plot(x,y,xlab="Independent",ylab="Dependent",main="Random Stuff")
points(x1,y1,col=2,pch=3)
points(x2,y2,col=4,pch=5)
legend(14,70,c("Original","one","two"),col=c(1,2,4),pch=c(1,3,5))
```

# Linear Least Squares Regression

```{r}
year <- c(2000 ,   2001  ,  2002  ,  2003 ,   2004)
rate <- c(9.34 ,   8.50  ,  7.62  ,  6.93  ,  6.60)

plot(year,rate,
     main="Commercial Banks Interest Rate for 4 Year Car Loan",
     sub="http://www.federalreserve.gov/releases/g19/20050805/")

cor(year,rate)

fit <- lm(rate ~ year)
fit

attributes(fit)

plot(year,fit$residuals)

summary(fit)
```

# Data Wrangling

```{r}
a <- data.frame(one=c( 0, 1, 2),two=c("a","a","b"))
b <- data.frame(one=c(10,11,12),two=c("c","c","d"))
a
b
```

```{r}
v <- rbind(a,b)
typeof(v)
v
```

```{r}
w <- cbind(a,b)
w
```

```{r}
names(w) = c("one","two","three","four")
w
```


```{r}
x <- list(a=rnorm(200,mean=1,sd=10),
            b=rexp(300,10.0),
            c=as.factor(c("a","b","b","b","c","c")))
lapply(x, summary)
```

```{r}
val <- data.frame(a=c(1,2,10,20,5,50),
                    b=as.factor(c("a","a","b","b","a","b")))
val
```

```{r}
result <- tapply(val$a,val$b,mean)
result
```

# OOP concepts in R 

if statments

```{r}
x = 0.1
if( x < 0.2)
  {
     x <- x + 1
     cat("increment that number!\n")
  }
x
```

```{r}
x = 1.0
if ( x < 0.2)
 {
    x <- x + 1
    cat("increment that number!\n")
 } else if ( x < 2.0)
 {
   x <- 2.0*x
   cat("not big enough!\n")
 } else
 {
    x <- x - 1
    cat("nah, make it smaller.\n");
 }
x
```

for statments 

```{r}

for (lupe in seq(0,1,by=0.3))
 {
    cat(lupe,"\n");
 }

x <- c(1,2,4,8,16)
for (loop in x)
 {
    cat("value of loop: ",loop,"\n");
 }
```

while statments

```{r}
lupe <- 1;
x <- 1
while(x < 4)
 {
    x <- rnorm(1,mean=2,sd=3)
    cat("trying this value: ",x," (",lupe," times in loop)\n");
    lupe <- lupe + 1
 }
```

Repeat statement

```{r}
repeat
  {
    x <- rnorm(1)
    if(x < -2.0) break
  }
x
```

Break and next statements

```{r}
x <- rnorm(5)
for(lupe in x)
 {
     if (lupe > 2.0)
         next

     if( (lupe<0.6) && (lupe > 0.5))
        break

    cat("The value of lupe is ",lupe,"\n");
 }
```

Switch statements

```{r}
x <- as.integer(2)
x
z = switch(x,1,2,3,4,5)
z
x <- 3.5
z = switch(x,1,2,3,4,5)
z
```

```{r}
x <- "sd"
z <- switch(x,"mean"=mean(y),"median"=median(y),"variance"=var(y),"sd"=sd(y))
z
```

Functions

```{r}
newDef <- function(a,b)
 {
     x = runif(10,a,b)
     mean(x)
}

newDef(-1,1)
```


References:
https://www.cyclismo.org/tutorial/R/

